---
title: "Using the earthquake capstone package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the earthquake capstone package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(ggplot2)
library(magrittr)
library(leaflet)
library(lubridate)
library(earthquakecapstone)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig_width = 8,
  fig_height = 5
)
```

This package is the capstone project for Coursera's
*Mastering Software Development in R* specialization consisting
of the courses

* [The R Programming Environment](https://www.coursera.org/learn/r-programming-environment?specialization=r),
* [Advanced R Programming](https://www.coursera.org/learn/advanced-r?specialization=r),
* [Building R Packages](https://www.coursera.org/learn/r-packages?specialization=r),
* [Building Data Visualization Tools](https://www.coursera.org/learn/r-data-visualization?specialization=r),
* [Mastering Software Development in R Capstone](https://www.coursera.org/learn/r-capstone?specialization=r)

The package provides functions to clean and visualize data from the
[NOAA Significant Earthquake Database](https://www.ngdc.noaa.gov/hazel/view/hazards/earthquake/search).


### Installation
Install the development version of this package using

```
devtools::install_github("lemonad/coursera-mastering-r-capstone")
```

### Dataset
The NOAA Significant Earthquake database can be made available by executing the
following

```{r}
api_uri <- "https://www.ngdc.noaa.gov/hazel/hazard-service/api/v1/earthquakes"
f <- jsonlite::read_json(api_uri, simplify = TRUE)
earthquakes <- dplyr::as_tibble(f$items)
```


### Cleaning
The API data has some quirks and can be made easier to work with by
using the function `eq_clean_data` which returns a dataframe that can be
subsequently filtered, etc.

```{r}
df <- earthquakes %>%
  eq_clean_data() %>%
  dplyr::filter(
    country %in% c("USA", "Mexico") & lubridate::year(date) >= 2000
  )
```

Note that the BCE events are removed from the results per the requirements
of using the class `Date` for the dates. This could be remedied by using the
package [gregorian](https://github.com/edgararuiz/gregorian).


### Timeline
Having cleaned the data, the events can now be plotted on a timeline
with optional label annotation. Below, the timeline is plotted with a
separate axis for each country and events are colored according to the
total number of deaths. The five highest magnitude events are annotated
with their corresponding location.

```{r}
df %>%
  ggplot2::ggplot(
    ggplot2::aes(x = date, y = country, colour = deathsTotal),
    alpha = 1
  ) +
  geom_timeline() +
  geom_timeline_label(
    ggplot2::aes(label = locationName, magnitude = eqMagnitude),
    n_max = 5
  )
```


### Map
The sites of the events can also be plotted on an interactive map. The sites
will show a tooltip containing details when clicked. Here, the helper
function `eq_create_label` is used to create the underlying html for the
tooltip.

```{r}
df %>%
  dplyr::mutate(popup_text = eq_create_label(.)) %>%
  eq_map(annot_col = "popup_text")
```
